---
title: "Introduction to AirMonitorPlots"
author: "Jonathan Callahan"
date: "October 11, 2023"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to AirMonitorPlots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  fig.width = 7,
  fig.height = 7
)
```

## Installation

Install from CRAN with:

`install.packages('AirMonitorPlots')`

Install the latest version from GitHub with:

`devtools::install_github('mazamascience/AirMonitorPlots')`

## Background

The **AirMonitorPlots** package provides plotting functionality to create 
production-ready plots for air quality monitoring data. It is designed to work
with the [AirMonitor](https://mazamascience.github.io/AirMonitor/) package, 
integrating the data with [ggplot2](https://ggplot2.tidyverse.org) plotting 
functions. High-level plotting functions make it easy for users to create beautiful 
plots of monitoring data. Since these functions are built on **ggplot2**, users 
familiar with **ggplot2** can easily  use the functions in this package to create 
custom plots. 

## Examples

The following examples use example _mts_monitor_ objects from the **AirMonitor** 
package.

### Timeseries

The default `Timeseries` plot is optimized for one monitor. 

```{r}
library(AirMonitor)
library(AirMonitorPlots)

monitor_ggTimeseries(
  AirMonitor::NW_Megafires, 
  startdate = 20150815,
  enddate = 20150825,
  id = "753019dc7f401908_410390101_03"
)
```

If more than one monitor is present, color will be mapped to deviceDeploymentID. 

```{r}
monitor_ggTimeseries(
  AirMonitor::NW_Megafires,
  startdate = 20150815,
  enddate = 20150820,
  id = c(
    "753019dc7f401908_410390101_03", 
    "3d1ac393596c511a_410390059_03", 
    "c69d9483556acb85_410391009_03",
    "eba98496f2495d0f_410390060_03"
  )
)
```

### Daily Barplot

The `DailyBarplot` shows daily averages for one monitor. 

```{r}
monitor_ggDailyBarplot(
  AirMonitor::Carmel_Valley,
  startdate = 20160801,
  enddate = 20160808
)
```

### Daily By Hour

The `dailyByHour` is designed specifically to show current data, with special 
lines for "today" and "yesterday". However, it can plot historical data, in 
which case `enddate` will be styled as "today".

```{r}
monitor_ggDailyByHour(
  AirMonitor::Carmel_Valley,
  startdate = 20160801, 
  enddate = 20160809
)
```

