---
title: "Introduction to AirMonitorPlots"
author: "Jonathan Callahan"
date: "April 09, 2022"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to AirMonitorPlots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  fig.width = 7,
  fig.height = 7
)
```

## Installation

_NOTE: This package is not yet available on CRAN._

Install the latest version from GitHub with:

`devtools::install_github('mazamascience/AirMonitorPlots')`

## Background

The **AirMonitorPlots`\** package provides plotting functionality to create 
production-ready plots for air quality monitoring data. It is designed to work
with the **AirMonitor** package, integrating the data with **ggplot2** plotting 
functions. High-level plotting functions make it easy for users to create beautiful 
plots of monitoring data. Since these functions are built on `ggplot2`, users 
familiar with `ggplot2` can easily  use the functions in this package to create 
custom plots. 

## Production-ready plots

High-level functions in this package make it easy to create beautiful,
publication-ready plots with very few lines of code.

The following plot types are provided:

**TODO**


## Examples

The following examples use example _mts_monitor_ objects from the **AirMonitor** 
package.

### Timeseries

The default `Timeseries` plot is optimized for one monitor. 

```{r}
library(AirMonitor)
library(AirMonitorPlots)

monitor_ggTimeseries_archival(
  AirMonitor::NW_Megafires, 
  startdate = 20150815,
  enddate = 20150825,
  deviceDeploymentIDs = "39c9acebbebc3d09_160050015"
)
```

If more than one monitor is present, color will be mapped to deviceDeploymentID. 

```{r}
monitor_ggTimeseries(
  AirMonitor::NW_Megafires,
  startdate = 20150815,
  enddate = 20150820,
  deviceDeploymentIDs = c(
    "fa8288b1da3b2a87_530630047", 
    "5b3acb7aa679dc14_530639997", 
    "abde4337eb9064e4_530639996"
  )
)
```

### Daily Barplot

The `DailyBarplot` shows daily averages for one monitor. 

```{r}
monitor_ggDailyBarplot(
  mts_monitor = AirMonitor::Carmel_Valley,
  startdate = 20160801,
  enddate = 20160808
)
```

### Daily By Hour

The `dailyByHour` is designed specifically to show current data, with special 
lines for "today" and "yesterday". However, it can plot historical data, using 
the enddate for "today". 

```{r}
monitor_ggDailyByHour(
  AirMonitor::Carmel_Valley,
  startdate = 20160801, 
  enddate = 20160809
)
```

