---
title: "Introduction to PWFSLSmokePlots"
author: "Helen Miller"
date: "Jan 17, 2019"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PWFSLSmokePlots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  fig.width = 7,
  fig.height = 7
)
```

`PWFSLSmokePlots` provides plotting functionality to create production-ready plots using air quality monitoring
data based on the data model from `PWFSLSmoke`. It builds on the `PWFLSmoke` package, integrating the data
with `ggplot2` plotting functions. High-level plotting functions make it easy for users to create beautiful 
plots of monitoring data. Since these functions are built on `ggplot2`, users familiar with `ggplot2` can easily 
use the functions in this package to create custom plots. 

## Production-ready plots

High-level functions in this package make it easy to create beautiful plots with very few lines of code. 

### Timeseries

The default Timeseries plot is optimized for one monitor. 

```{r}
library(PWFSLSmoke)
library(PWFSLSmokePlots)
monitor_ggTimeseries(PWFSLSmoke::Northwest_Megafires, 
                     monitorIDs = "160050015_01",
                     startdate = 20150815,
                     enddate = 20150825)
```

If more than one monitor is present, color will be mapped to monitorID. 

```{r}
monitor_ggTimeseries(PWFSLSmoke::Northwest_Megafires,
                     monitorIDs = c("530630047_01", 
                                    "530639997_01", 
                                    "530639996_01"),
                     startdate = 20150815,
                     enddate = 20150820)
```

### Daily Barplot

Daily Barplot can show daily averages for one monitor. 

```{r}
monitor_ggDailyBarplot(ws_monitor = PWFSLSmoke::Carmel_Valley,
                       startdate = 20160801,
                       enddate = 20160807)
```

It can also show aggregated daily averages for many monitors.

```{r}
monitor_ggDailyBarplot(PWFSLSmoke::Northwest_Megafires,
                       startdate = 20150820, 
                       enddate = 20150825)
```

### Daily By Hour

The `dailyByHour` plot is designed to view current data for only one monitor. It can show daily trends in air quality.

```{r}
latest <- airnow_loadLatest()
monitor_ggDailyByHour(latest,
                      monitorID = "060631010_01")
```


## Extending `ggplot2`

High-level plotting functions create plots by adding custom layers and styling, returing a `ggplot` object. 
The package extends `ggplot2`, including custom geoms, stats, and themes, as well as some `custom_*` functions
which combine layers and other styling options to make consistent and appropriately styled plots. A plot can
be instantiated with a `ggplot_*` function. Choose the correct function, depending on the type of plot. Then,
layers and styling can be added with functions from this package or `ggplot2` functions. This section demonstrates
how these pieces work together, and shows some examples of the kinds of custom plots that can be made using this functionality. 

### pm25Timeseries plots

The most basic type of plot for PWFSL monitoring data is timeseries plots: plots that have time on the x-axis
and PM2.5 on the y-axis. Timeseries plots should start with a call to `ggplot_pm25Timeseries`, which accepts
either a `ws_monitor` object, or a `ws_tidy` object, created from `monitor_toTidy(ws_monitor)`. This 
instantiates the plot mapping the data to the correct axes, and adds some custom styling and scaling. 


#### Many Monitors

```{r}
nmPlot <- ggplot_pm25Timeseries(Northwest_Megafires,
                           startdate = 20150815,
                           enddate = 20150825) + 
  ggtitle("2015 Northwest Megafires")

nmPlot
```

Once the plot is initiated, you can add layers.

```{r}
nmPlot + 
  geom_pm25Points(shape = "square", alpha = .1) + 
  stat_dailyAQILevel(alpha = .5) + 
  scale_y_continuous(limits = c(0, 500)) +
  custom_aqiStackedBar(width = 0.01)
```

#### Several Monitors

```{r}
# Subset out several monitors near Spokane
monitorIDs <- c("530630047_01", "530639997_01", "530639996_01")
spokane <- monitor_subset(Northwest_Megafires, monitorIDs = monitorIDs)

spokanePlot <- ggplot_pm25Timeseries(spokane,
                                     startdate = 20150815,
                                     enddate = 20150820) +
  ggtitle("Smoke Monitors in Spokane, 2015")

spokanePlot + 
  geom_pm25Points(aes(color = monitorID)) + 
  stat_nowcast(aes(color = monitorID))
spokanePlot + 
  stat_dailyAQILevel(timezone = "America/Los_Angeles") + 
  facet_grid(rows = vars(monitorID))
spokanePlot + 
  stat_AQILevel(color = NA) + 
  stat_dailyAQILevel(alpha = .5, timezone = "America/Los_Angeles") + 
  facet_grid(rows = vars(monitorID)) 
spokanePlot + 
  custom_aqiStackedBar(width = 1, alpha = .3) + 
  stat_nowcast(aes(linetype = monitorID))
spokanePlot + 
  custom_aqiLines() + 
  stat_nowcast(aes(linetype = monitorID)) + 
  custom_pm25TimeseriesScales(spokane, startdate = 20150810, enddate = 20150820) 

```

### pm25Diurnal plots

There are also some functions designed to create diurnal plots -- plots designed to show daily trends. They should start with a call to 
`ggplot_pm25Diurnal`. 

#### One monitor
This type of plot is best suited for exploring one monitor, although a savvy `ggplot2` user could certainly create useful and attractive 
diurnal plots representing multiple monitors. Like `ggplot_pm25Timeseries`, `ggplot_pm25Diurnal` simply instantiates the plot without adding
any layers yet. By default, it includes day/night shading, although this can be turned off by specifying `shadedNight = FALSE`. 

```{r}
monitorID <- "530630047_01"
monroeStMonitor <- monitor_subset(Northwest_Megafires, monitorIDs = monitorID)
ggplot_pm25Diurnal(monroeStMonitor)
```

After instantiating the plot, you can add layers. `stat_meanByHour` is a stat specifically designed for this type of plot, which 
calculates the mean y-value for each x-value. You can also use the 'group' aesthetic to tell a stat to do calculations by a group
such as hour or day. There is also an `offsetBreaks` option, which, if `TRUE`, will offset the x-axis guides and ticks by 0.5, which 
is useful for plots like boxplots where you might want the boxes to be between the hour ticks instead of lined up with them. For example: 

```{r}
# offsetBreaks = FALSE. Notice that bars cover the x-axis guides
ggplot_pm25Diurnal(PWFSLSmoke::Carmel_Valley, 
                   startdate = 20160801, 
                   enddate = 20160810)  + 
  stat_meanByHour(output = "AQIColors")

# offsetBreaks = TRUE
ggplot_pm25Diurnal(PWFSLSmoke::Carmel_Valley,
                   startdate = 20160801, 
                   enddate = 20160810,
                   offsetBreaks = TRUE) + 
  stat_meanByHour(outpu = "AQIColors")

  
```

Other examples
```{r}

ggplot_pm25Diurnal(PWFSLSmoke::Carmel_Valley,
                   startdate = 20160801, 
                   enddate = 20160810,
                   offsetBreaks = TRUE) +
  geom_boxplot(aes(group = hour)) +
  custom_aqiStackedBar()

ggplot_pm25Diurnal(monroeStMonitor,
                   startdate = 20150815,
                   enddate = 20150820) + 
  stat_nowcast(aes(color = day), geom = "path", size = 1) + 
  stat_meanByHour(geom = "line", size = 3) 

ggplot_pm25Diurnal(spokane, 
                   startdate = 20150815,
                   enddate = 20150820) +
  stat_meanByHour(aes(color = monitorID), geom = "path")

# "clock" plot

ggplot_pm25Diurnal(monroeStMonitor,
                   startdate = 20150815,
                   enddate = 20150820,
                   offsetBreaks = TRUE,
                   xexp = c(1/41, 1/41)) + 
  coord_polar() + 
  stat_meanByHour(aes(y = 75), input = "pm25", output = "AQIColors", width = 1, color = adjustcolor("black", .2))
```

